### 2021-10-22 paper.Rmd
some discussion, abstract

    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while aiming for two objectives: (a) high between-group similarity and (b) high within-group heterogeneity. For some anticlustering formulations, these two objectives coincide, i.e., high between-group similarity is achieved when within-group heterogeneity is maximized. Other anticlustering criteria focus on one of the two aspects. The present paper focuces on maximizing between-group similarity by introducing a novel bicriterion objective for anticlustering. *K-plus anticlustering* is an extension of k-means anticlustering, which minimizes the difference between the mean of numeric attributes across clusters. The k-plus objective extends k-means by adding a term that aims to optimize similarity with regard to the variance of the attributes data across clusters. It is shown that the combined bicriterion objective can be reduced to an augmentation of the input data, leaving the original k-means criterion unchanged: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original values. A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion. The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity. K-plus anticlustering is easily accessible to researchers and practitioners: the method has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

In this paper, I presented k-plus anticlustering, a method to divide a data set into similar subgroups. The k-plus criterion extends the classical k-means criterion by incorporating a term that represents how groups differ with regard to their variances. The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable can be computed as the squared deviation between the original values and the overall mean of the original variable. A simulation study and examples on norming data indicate that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.

K-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups. For small samples sizes (or large $K$), k-plus anticlustering may return slightly less-than-optimal results in the eye of the user, but in this case it is no problem to return to k-means and only focus on the means. There is no anticluster hacking; as opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering. Users may simply choose the results they like best. As a starter, I strongly recommend trying out k-plus anticlustering when creating groups that aim for high between-group similarity.

>>>>

    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while aiming for two objectives: (a) high between-group similarity and (b) high within-group heterogeneity. For some anticlustering formulations, these two objectives coincide, i.e., high between-group similarity is achieved when within-group heterogeneity is maximized. The present paper focuces on maximizing between-group similarity by introducing the k-plus objective, a novel bicriterion for anticlustering. K-plus anticlustering is an extension of classical k-means anticlustering. While k-means anticlustering focuses on minimizing between-group-differences with regard to the means of attribute values, k-plus anticlustering includes a term to optimize similarity with regard to the variance at the same time. The k-plus bicriterion can be reduced to an augmentation of the input data: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original variable. Thus, k-plus leaves the original k-means criterion unchanged and only differs in the input data. A computer simulation and practical examples on real norming data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion. The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity. K-plus anticlustering is easily accessible as it has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

In this paper, I presented the k-plus criterion for anticlustering. Anticlustering methods are used to divide sets of elements into subgroups while striving for high between-group similarity or high within-group heterogeneity. The k-plus criterion represents between-group similarity as the discrepancy between the mean and variance of attribute values. These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided. The k-plus criterion extends the classical k-means criterion---which represents how similar groups are with regard to their means---by incorporating a term that represents how groups differ with regard to their variances. The combined bicriterion objective could be reduced to an augmentation of the input data; for each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable. Thus, the original k-means criterion remains basically unchanged. A simulation study and examples on norming data indicated that optimizing the new k-plus objective is well-suited for minimizing differences with regard to means and variances between groups.

This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time. However, for small group sizes (small $N$ and / or large $K$) k-plus anticlustering may return slightly less-than-optimal results in the eye of the user. In this case, it is not a problem to return to k-means that only focuses on the means. As opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering. There is no anticluster hacking. Users may simply choose the results they like best. As a default, it is recommended to apply out k-plus anticlustering when creating groups that aim for high between-group similarity.
